---
title: "Diversity Outbred Micxe"
author: "Brian S. Yandell"
date: "12/16/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a study of Diversity Outbred (DO) mice derived from 8 founder strains.

Data: <https://github.com/rqtl/qtl2data/tree/master/DO_Gatti2014>  
Reference: Gatti DM, Svenson KL, Shabalin A, Wu L-Y, Valdar W, Simecek P, Goodwin N, Cheng R, Pomp D, Palmer A, Chesler EJ, Broman KW, Churchill GA (2014) Quantitative trait locus mapping methods for Diversity Outbred mice. G3 4:1623-1633 [doi:10.1534/g3.114.013748](http://dx.doi.org/10.1534/g3.114.013748)

Here are some useful links:

- <http://rqtl.org> (lots of useful stuff here)
    + A brief tour of R/qtl: [[pdf](http://rqtl.org/tutorials/rqtltour.pdf) | [code](http://rqtl.org/tutorials/rqtltour.R)]
    + A shorter tour of R/qtl: [[pdf](http://rqtl.org/tutorials/rqtltour2.pdf) | [code](http://rqtl.org/tutorials/rqtltour2.R)]
    + [R/qtl sample data formats](http://rqtl.org/sampledata/)
    + [A Guide to QTL Mapping with R/qtl by Karl W. Broman & Śaunak Sen](http://www.rqtl.org/book/)
- <http://kbroman.org/qtl2/>
    + See Installation Instructions
    + [R/qtl User Guide](http://kbroman.org/qtl2/assets/vignettes/user_guide.html)
    + [R/qtl2 input file format](http://kbroman.org/qtl2/assets/vignettes/input_files.html).

Files for DO mice are in the [do](do) folder. These include:

- [kinship.csv](do/kinship.csv): list of kinship matrices by chr (20Mb)
- [do_covar.csv](do/covar.csv): covariate matrix (11Kb)
- [probs.csv](do/probs.csv): genotype probabilities object (2.1Gb)
- `pheno_*.csv`: phenotype data frames for `clin`, `otu`, `otufam` (~1.5Mb)
- [pmap.csv](do/pmap.csv): list of physical maps by chr (1Mb)
- [ccfoundersnps.sqlite](do/ccfoundersnps.sqlite): SQLite SNP database (2.9Gb)

## Setup

First we set up needed packages, or libraries.

```{r message=FALSE}
library(qtl2scan)
library(qtl2plot)
library(qtl2geno)
library(dplyr)
```

Could read in data over the internet. We do it here with local copy.

```{r eval=FALSE}
## File address across internet.
file <- paste0("https://raw.githubusercontent.com/rqtl/",
               "qtl2data/master/DO_Gatti2014/do.zip")
```

```{r}
## Local file (if saved on your machine)
file <- paste0("do.zip")
```

Now read the cross data

```{r}
do <- read_cross2(file)
```

```{r}
dimnames(do$pheno)[[2]]
```

## Genome scan

Calculate genotype probabilities. This takes some time and may create a huge file.
The original ZIP file with all data was 1.5Mb.
The calculated genotype probability file is 1.1Gb with all 36 "diplotype" probabilites.
We reduce to the 8 allele "haplotypes", which is only 246Mb.
We also put the physical map into the probs object.
We save both objects, as they took some time to create.


```{r}
if(file.exists("do_probs.rds")) {
  do_probs <- readRDS("do_probs.rds")
} else {
  probs <- calc_genoprob(do, step=1, err=0.002)
  saveRDS(probs, "do_36_probs.rds")
  do_probs <- genoprob_to_alleleprob(probs)
  do_probs$map <- do$pmap
  saveRDS(do_probs, "do_probs.rds")
}
```

Calculate kinship matrices. This takes a short while, and is 78Mb.

```{r}
if(file.exists("do_kinship.rds")) {
  do_kinship <- readRDS("do_kinship.rds")
} else {
  do_kinship <- calc_kinship(do_probs, "loco")
  saveRDS(do_kinship, "do_kinship.rds")
}
```

```{r}
pheno_scan <- scan1(do_probs, 
                   do$pheno[,2],
                   do_kinship)
```

```{r}
plot(pheno_scan)
```

```{r}
(peak <- max(pheno_scan))
```

```{r}
plot(pheno_scan, chr="1")
```

```{r}
pheno_1 <- scan1(subset(do_probs, chr="1"),
                   do$pheno[,2],
                   do_kinship[["1"]])
```

```{r}
plot(pheno_1)
```

## Allele Scan

```{r}
eff_scan <- scan1coef(subset(do_probs, chr = "1"), 
                      do$pheno[,2],
                      do_kinship[["1"]])
```

```{r}
plot_coefCC(eff_scan)
abline(v=peak$pos, lty=2, lwd=2)
```


## SNP Scan

In order to reproduce the SNP scan, we need to get the SNP information.
These are MUGA SNPs, with information at <ftp://ftp.jax.org/MUGA/>.
You must download <ftp://ftp.jax.org/MUGA/DO_mice/muga_snps.Rdata>

Problem is that the `muga_snps.Rdata` file does not have SDP,
and that is not easily extracted from genotype probabilities.

```{r}
load("muga_snps.Rdata")
window_Mbp <- 3
```


```{r}
snpinfo <- muga_snps %>%
  filter(chr == "1",
         pos >= peak$pos - 3,
         pos <= peak$pos + 3)
dim(snpinfo)
```


```{r eval=FALSE}
snpprobs <- genoprob_to_snpprob(subset(probs, chr = "1"),
                                snpinfo)
```

```{r}
pheno_snpscan_chr <- scan1(snpprobs, 
                   phe_df[, pheno_names, drop = FALSE],
                   K[[chr_id]],
                   covar[, covar_names])
```

```{r}
plot(pheno_snpscan_chr, drop.hilit = 1.5)
```

```{r}
pheno_topsnps <- top_snps(pheno_snpscan_chr, drop = 1.5)
pheno_topsnps
```

